\begin{acknowledgments}
I am deeply grateful to my advisor Kathryn McKinley for guiding me through
grad school with utmost interest and dedication. She was flexible enough to
let me do what I wanted, even if it was outside her area. In the days
before my first paper submission, she gave me undivided attention and
worked through the night until we submitted the paper.  At times it seemed
like she cared more about my success and well-being than I did myself.  I
wouldn't have gotten this far without her unflinching positive attitude and
encouragement. As I leave grad school, inspired by Kathryn as a researcher
and a human being, I will strive to always give my best to those around me.

Over the past three years, Michael Hicks' meticulous attitude towards
research has been a positive influence. Always prompt to reply to e-mails,
he invariably shot down most of my arguments and imparted some of his vast
knowledge in the field of dynamic software updating. I thank him for all
his advice and for introducing us to scrum for research~\cite{scrum}.

Steve Blackburn has always been eager to offer technical advice and
guidance. I thank Steve, Keshav Pingali, and Calvin Lin for their
insightful questions and helpful feedback.

My research has benefitted enormously from my interactions with Mike Bond.
I thank him for brainstorming engaging research problems as well as for
patiently answering technical questions. For their personal and technical
support, I thank members of the Speedway research group: Jenn Sartor, Bert
Maher, Byeongcheol Lee, Katie Coons, Dimitrios Prountzos, Ivan Jibaja, Alex
Loh, and Na Meng.

This research would not have been possible without \RVM. I thank \RVM's
developers for maintaining an excellent research infrastructure, and the
community, especially Dave Grove and Filip Pizlo, for helpful advice.

I am thankful to Gem Naivar, Tom Horn, Lydia Griffith, and Gloria Ramirez
for all their guidance and help. My thanks are also due to UT and NSF for
directly and indirectly funding my research.

Among my friends, I am indebted to Smriti for her patient listening in
times good and bad, and my roommates, Sibi and Ramtilak, for all their
support. I thank Smita, Urmila, and other volunteers of Vibha for all the
good times and Vishaal, Harish, and Rohan for numerous music sessions that
helped keep me sane. My time in Austin was made memorable by several nice
folks. I thank them all.

I wouldn't have reached here without my family. My parents Swarnam and
Subramanian showered me with love and affection, gave me a great education,
and instilled in me the importance of sincerity and hard work. My brother
Shriram has been my best cheerleader. To my family, I dedicate this
dissertation.
\end{acknowledgments}
