% \section{Other proposals}
% 
% This part is not well written. Appreciate any suggestions.

\section{Updates in a persistent object store}
\label{sec:boyapati}

Boyapati et al.~\cite{boyapati03lazy} support dynamic updates to classes
kept in a \emph{persistent object store} (POS).  While the setting is
different, their basic update model, and in particular their notion of
object transformer function, is similar to ours.  In their system,
programmers manually write an object transformer that they view as a method
on the old version of the updated class, i.e., the transformer method is
type-safe with respect to the old class.  In \JV, object transformers may
access the \emph{new} versions of objects pointed to from the old class.
Instead, Boyapati et al.'s transformers may access the \emph{old} versions.
To implement this model, they rely on \emph{encapsulation} based on
\emph{ownership types}: if an object $a$ of class $A$ has an ``owned''
field pointing to an object $b$ of class $B$, then only $a$ can point to
(and access) $b$.  Encapsulation thus ensures the system will always
transform $a$ before $b$, which makes the transformation algorithm more
efficient.  They rely on the programmer to enforce encapsulation, and
describe how the compiler could automate language support for encapsulation
in a non-standard type system.  \JV takes the opposite tack of forcing old
object fields to point to up-to-date objects, and thus requires no special
language support.  Moreover, \JV's model follows that of earlier
work~\cite{k42usenix, neamtiu06dsu, neamtiu09stump, upstare} which has
proven its effectiveness on a half-dozen realistic applications across
several years' worth of releases.
% However, further research to understand
% the costs and benefits of the two updating models would be useful.

Boyapati et al. also differs from \JV in that, like JDRUMS and DVM, updates
are applied incrementally as objects are accessed following an update
rather than all at once using a stop-the-world GC\@. This incremental cost
is more natural in a POS since indirection is already required to access
external objects. The POS model also permits programmers to specify ACID
transaction boundaries, which can help ensure that updates are applied
consistently and safely.  In contrast, our work focuses on supporting
dynamic upgrades in a high-performance VM for Java, and thus many of the
issues we consider---reaching a safe point via return barriers and OSR, and
coexisting with the JIT compiler---are the unique contributions of our
work.
