\documentclass[natbib]{sigplanconf}

\usepackage{color}
\usepackage{mathptmx}
\usepackage[scaled=.92]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{listings}
\lstset{language=Java,basicstyle=\small\sffamily,columns=flexible,numberstyle=\em\scriptsize,frame=none,mathescape=true}
\newcommand{\code}[1]{\lstinline|#1|\xspace}
\newcommand{\javac}{javac\xspace} % This package lets you punctuate \javac normally and get good spacing, e.g., \javac.  gives you: javac.
\usepackage{url}  % particularly useful for URLs in bib entries
\input{smagill-base}

\newcommand{\TOSAcronym}{\emph Targeted Object Synthesis (TOS)\xspace} 
\newcommand{\TOS}{TOS\xspace} 

\newcommand{\stephen}[1]{\textcolor{blue}{Stephen: #1}}
\newcommand{\sbm}[1]{\textcolor{blue}{Stephen: #1}}
\newcommand{\suriya}[1]{\textcolor{blue}{Suriya: #1}}
\newcommand{\kathryn}[1]{\textcolor{blue}{Kathryn: #1}}
\newcommand{\mike}[1]{\textcolor{blue}{Mike: #1}}

\begin{document}

\conferenceinfo{Mystery'11,} {January 1, 2011, Paris, TX.}
\CopyrightYear{2011}
\copyrightdata{2011}


\title{Automating Object Transformations for Dynamic Software Updating
%   \thanks{COMMENT OUT FOR SUBMISSION, UNCOMMENT FOR final version. Some of these may or may not apply to your work.  Make Kathryn check.
%     This work is supported by NSF SHF-0910818, NSF CSR-0917191, NSF
%     CCF-0811524, NSF CNS-0719966, NSF CCF-0429859, Intel, IBM, CISCO,
%     Google, and Microsoft.  Any opinions, findings and conclusions
%     expressed herein are the authors' and do not necessarily reflect
%     those of the sponsors.}
    }

% 2008: NSF EIA-0303609, DARPA F33615-03-C-4106,


\authorinfo{Anonymous}{}{}
%% \authorinfo{Stephen Magill, Michael Hicks}
%%            {University of Maryland, College Park}
%%            {smagill@cs.umd.edu}

%% \authorinfo{Suriya Subramanian}
%%            {Intel Corporation}
%%            {suriya@gmail.com}

%% \authorinfo{Kathryn S. McKinley}
%%            {Microsoft Research}
%%            {mckinley@cs.utexas.edu}

\maketitle

\input{abstract}

% {\scriptsize
% \category{D.3.4}{Programming Languages}{Processors}[Memory management (garbage collection); Optimization] 
% \terms
% Experimentation, Languages, Performance, Measurement
% \keywords
% Heap
% }

\input{defs.tex}

\input intro

\input overview

\input related

\section{Example}

\input matching

\section{Synthesis Language}

\section{Using Synthesis to Guide Matching}

This is something Mike and I talked about at POPL.  It seems like a
good idea that could help us match objects even when there is not a
clear set of ``key fields''.  The issue is complexity.  Suppose we
have $M$ old-version objects and $N$ new-version objects (both of the
same class).  There are $M \times N$ possible pairings.  Of course we
are already doing $M \times N$ edit distance computations in the
matching phase, but these comparisons are likely much faster than
synthesis will be.

One hybrid approach would be to use edit-distance-based matching and
then use synthesis to fine-tune the matching in cases that are not
clear-cut.  So if $o_1$ is edit distance 4 from $o_2$ and edit
distance 5 from $o_3$, then maybe use synthesis to help determine
which pairing should be chosen.

We could also use synthesis to check whether it is a good idea to
introduce a key field.  If there are no natural key fields, but we
think that allocation order might be a good matching heuristic, we
could try it and see if synthesis succeeds.

It seems that when using synthesis as a ``sanity check'' for the
matching step, it is better to be doing synthesis for a less
expressive language.  The more expressive the language is, the less
likely synthesis is to fail, making it less useful as a filter for
matching.  This is just intuition though and could be totally wrong.

\input{experiments}

% \input{beef}
% \input{results.tex}

\input{conclusion.tex}


\bibliographystyle{abbrvnat}
% \renewcommand{\bibfont}{\footnotesize} % <--- change bib font size here
% \setlength{\bibsep}{0.5ex}             % <--- change space between bib entries here
\bibliography{paper}

\input oldtext 

\input conflicts

\end{document}
