\section{Experience}
\label{sec:experience}

To evaluate \DSU, we used it to update three open-source servers written
in Java: the Jetty webserver\footnote{\url{http://www.mortbay.org}},
JavaEmailServer,\footnote{\url{http://www.ericdaugherty.com/java/mailserver/}}
an SMTP and POP e-mail server, and CrossFTP server.\footnote{\url{http://www.crossftp.com/}}
These programs belong to a class that
should benefit from DSU because they typically run continuously. DSU
would enable deployments to incorporate bug fixes or add new features
without having to halt currently-running sessions.  

We explored
updates corresponding to releases made over roughly one to two years
of each program's lifetime.  Of the 22 updates we considered, \DSU{} could
support 20 of them---the two updates we could not apply changed
classes with infinitely-running methods, and thus no safe point could
be reached.  To our knowledge, no existing DSU system
for Java could support all these updates; indeed, previous systems
with simple support for updating method bodies would be able to handle only 9 of the 22 updates.  Although \DSU{} cannot support every
update, it is the first DSU system for Java
that has been shown to support changes to realistic programs as they
occur in practice over a long period of time.

In the rest of this section, we first examine the performance impact of
\DSU{}, and then look at updates to each of the three applications in
detail.

\input{experience/performance}
\input{experience/experience-jetty}
\input{experience/experience-jes}
\input{experience/experience-crossftp}
