% \input{experience/performance-jetty-table}
% \input{experience/performance-jetty-graph}
% \input{experience/performance-microbench-table}

\paragraph{Jetty performance.}
To see the effect of updating on application performance, we measured Jetty
under various configurations using httperf, a webserver
benchmarking tool.\footnote{\url{http://www.hpl.hp.com/research/linux/httperf}}  We used httperf to issue roughly 800 new connection
requests/second, which we observed to be Jetty's saturation rate.
Each connection makes 5 serial requests for a 40 Kbyte file. Httperf
reports average throughput and average per-request latency over a 60 second period. We
ran this experiment 21 times and report the median and
quartiles of the throughput and latency reports. With 21 runs, the range between
the quartiles serves as a 98\% conﬁdence interval~\cite{PrattGibbons81}.
In order to eliminate network traffic
effects, we ran the server on two cores of a quad-core machine and the
client on another core.

Figure~\ref{fig:jetty} shows our results in tabular form and plotted graphically.  The second and third columns
of the table
report the median throughput and the range between the two quartiles.
The third column and fourth column report the median latency and the
inter-quartile range.  The
first and seconds rows illustrate the performance of Jetty version 5.1.6
running on stock \JikesRVM{} and \DSU, respectively. The third row shows
the performance on \DSU{} of Jetty 5.1.6 dynamically updated from version
5.1.5 prior to the start of the experiment.
% \todo{Find a way to introduce the graph} Figure~\ref{fig:jetty} plots the results shown in the table.
The performance of the two \DSU{} configurations
is essentially identical: the two configurations' corresponding
inter-quartile ranges largely overlap.
The performance of \DSU{} is also quite similar to the performance of
stock \JikesRVM.  There are many small differences between \DSU{} and the stock
implementation that change VM code size, code layout, and garbage
collection behavior.  These differences may impact performance
directly and they may indirectly affect other elements of the VM,
such as the timing of garbage collections and JIT
optimizations (such indirect effects make VMs notoriously difficult to
benchmark~\cite{dacapo-cacm}).


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../pldi64"
%%% End: 
